Non-word repetition in Yélî Dnye: A small-scale study

### 2019-10-30 manuscript.Rmd
ana in progress

@@

The child was donned a headset (xx @mc please fill in for most of the children, xx @ac fill in for the rest). For most children, the headset could not stay comfortably on the child's head, and thus it was placed on the child's shoulders, with the microphone carefully placed close to the child's mouth. A local informant sat next to the child, to would provide the instructions and, if needed, coach the child to make sure, using the three practice items as well as real words, that they understood that the task was to repeat the items precisely without changing anything. An experimenter delivered the elicitation stimuli to the local informant and the child over headhphones.

@@

NWR studies vary in whether children are provided with several opportunities to hear and say the item. To have a fixed and clear procedure, we decided that items other than the inital three training ones would not be repeated unless the child made an attempt to produce them. If this attempt was judged correct by the local informant, then the experimenter would move on to the next item (whispering this over a headset that was recording a separate track). If the local informant heard a deviation, she indicated to the experimenter that the item needed to be repeated, and up to 5 attempts were allowed.

@@

```{r wrangling}
#@mc look here - we probably need to read the table in a special way to get all the accents etc. right
read.table("NWR-transcription.txt",header=T,enc="UTF-16le")->trnsc
read.csv("NWR-transcription.csv",sep=";",enc="UTF-8")->trnsc
trnsc$tokid=paste(trnsc$item,trnsc$token)
#task 4 are probably the items with no target; there are 1488 lines of paired data

@@

@@

```{r phonoana}

@@

## TODO
# make all vowels in mispronunciation & target short (so we don't penalize length errors)
# if correct=0 but orthotarget=mispronunciation (probably due to length errors) then change correct to 1

@@

trnsc$mp_uni=trnsc$mispronunciation #start from ortho
trnsc$target_uni=trnsc$target.ortho #start from ortho

@@

#log non-fluent speech
trnsc$nonfluent<-NA
trnsc$nonfluent[grep("-",trnsc$mp_uni,fixed=T)]<-1

#trnsc$additions=
```

```{r}
#create a subset with just the first presentations -- this is what is going to be used for the quantitative analyses
firstp=subset(trnsc,rep==1 & !is.na(rep))
```

- What is the overall repetition accuracy (whole word, phoneme based, distance)?

- How does this change as a function of item complexity (number of syllablex, sound complexity)?

- How frequent are errors that result in real words? Is that a function of item complexity? What are some common patterns of phonological errors?

>>>>

```
## TODO

- create stimuli table that has complexity & length
- integrate demo data & stim table in final data
- add analysis on the basis of complexity & length
- add analysis on the basis of demo chars
- fix the back-transformation
- add summary of types of error: %semantic, %deletion (which sounds?) %subst (which are the most common sounds) 
- maybe add to the last ana the prevalence (ie out of all the gh which proportion get transformed)
- discussion

@@

The child was donned a headset (xx @mc please fill in for most of the children, SHURE WH20 XLR headset with a dynamic microphone for the rest), recorded into the left channel into a Tascam DR40x digital recorder. For most children, the headset could not stay comfortably on the child's head, and thus it was placed on the child's shoulders, with the microphone carefully placed close to the child's mouth. A local informant sat next to the child, to would provide the instructions and, if needed, coach the child to make sure, using the three practice items as well as real words, that they understood that the task was to repeat the items precisely without changing anything. An experimenter delivered the elicitation stimuli to the local informant and the child over headphones.

@@

NWR studies vary in whether children are provided with several opportunities to hear and say the item. To have a fixed and clear procedure, we decided that items other than the inital three training ones would not be repeated unless the child made an attempt to produce them. If this attempt was judged correct by the local informant, then the experimenter would move on to the next item (whispering this over a separate headset that was recording onto the right track of the same Tascam). If the local informant heard a deviation, she indicated to the experimenter that the item needed to be repeated, and up to 5 attempts were allowed.

@@

```{r wrangling, eval=T}
read.delim("NWR-transcription.txt", encoding="UTF-8")->trnsc
trnsc$tokid=paste(trnsc$item,trnsc$token)
#task 4 are probably the items with no target; there are 1488 lines of paired data

@@

##TODO integrate demo data

@@

```{r phonoana, eval=T}

@@

#log non-fluent speech, then get rid of that marker
trnsc$nonfluent<-NA
trnsc$nonfluent[grep("-",trnsc$mp_uni,fixed=T)]<-1
trnsc$mispronunciation=gsub("-","",trnsc$mispronunciation)

# make all vowels in mispronunciation & target short (so we don't penalize length errors)
trnsc$mp=gsub("([aeiouêâéáóî])\\1+", "\\1", trnsc$mispronunciation)
trnsc$target.ortho=gsub("([aeiouêâéáóî])\\1+", "\\1", trnsc$target.ortho)

# if correct=0 but orthotarget=mispronunciation (probably due to length errors) then change correct to 1 
#sum(trnsc$correct==0 & trnsc$target.ortho == trnsc$mp & !is.na(trnsc$mp)) #this only happens twice
trnsc$correct[trnsc$correct==0 & trnsc$target.ortho == trnsc$mp & !is.na(trnsc$mp)]<-1

@@

trnsc$mp_uni=trnsc$mp #start from ortho
trnsc$target_uni=trnsc$target.ortho #start from ortho

@@

#calculate Levinshtein's distances 
trnsc$ins=trnsc$del=trnsc$sub=trnsc$nchar=trnsc$trafos=trnsc$tarlen=NA
substitution_bank=NULL
deletion_bank=NULL
for(i in grep("0",trnsc$correct)){
 #lev
  lev=adist(trnsc[i,c("mp_uni")],trnsc[i,c("target_uni")],count=T)
  trnsc$ins[i]<-attributes(lev)$counts[,,"ins"]
  trnsc$del[i]<-attributes(lev)$counts[,,"del"]
  trnsc$sub[i]<-attributes(lev)$counts[,,"sub"]
  trnsc$trafos[i]=attributes(lev)$trafos
  trnsc$nchar[i]=nchar(attributes(lev)$trafos)
  trnsc$tarlen[i]=nchar(trnsc$target.ortho[i])
  sublist=which(strsplit(attributes(lev)$trafos, "")[[1]]=="S")
  substitution_bank=rbind(substitution_bank,cbind(strsplit(trnsc[i,c("target_uni")],"")[[1]][sublist],strsplit(trnsc[i,c("mp_uni")],"")[[1]][sublist]))
  dellist=which(strsplit(attributes(lev)$trafos, "")[[1]]=="D")
  deletion_bank=rbind(deletion_bank,cbind(strsplit(trnsc[i,c("target_uni")],"")[[1]][dellist]))
}

trnsc$ld=rowSums(trnsc[,c("ins","del","sub")])
trnsc$nld=trnsc$ld/trnsc$nchar
trnsc$nld[is.na(trnsc$nld)]<-0

#this is not working
#trnsc$pc=nchar(gsub("SDI","",trnsc$trafos))/trnsc$nchar
#trnsc$pc[is.na(trnsc$pc)]<-1

write.table(trnsc,"final_data.txt",row.names=F,sep="\t")

#TODO this is not working correctly
# for(i in dim(correspondances)[1]:1) { #transform into orthography
#   substitution_bank=gsub(correspondances[i,2],correspondances[i,1],substitution_bank)
#   deletion_bank=gsub(correspondances[i,2],correspondances[i,1],deletion_bank)
# }

substitution_bank[order(substitution_bank[,1]),]->substitution_bank
write.table(substitution_bank,"substitution_bank.txt",row.names=F,sep="\t")

sort(deletion_bank)->deletion_bank
write.table(deletion_bank,"deletion_bank.txt",row.names=F,sep="\t")
```

```{r readin}
read.table("final_data.txt",header=T,encoding="UTF-8")->trnsc
#create a subset with just the first presentations -- this is what is going to be used for the quantitative analyses
firstp=subset(trnsc,rep==1 & !is.na(rep))
```

Taking into account only the first attempts, the overall repetition accuracy measured in whole words is `r round(mean(firstp$cor.online)*100)`% based on the online coding (i.e., whether the native experimenter asked for the target to be re-played), and `r round(mean(firstp$correct)*100)`% when the offline coding is considered. On average, the phoneme-based normalized Levenshtein distance was `r round(mean(firstp$nld)*100)`%, meaning that about 20% of phonemes were substituted, inserted, or deleted.

We then inspected whether accuracy varied as a function of word length.

```{r figlength,eval=F}
#not working
aggregate(firstp$correct,list(firstp$tarlen,firstp$id),mean)
plot(firstp$correct~firstp$tarlen)

```
- How frequent are errors that result in real words? Is that a function of item complexity? What are some common patterns of phonological errors?

### 2019-10-30 manuscript.tex
ana in progress

@@

Although infants begin to learn about their native language's phonology within the first year, many studies suggest that in perception and production, in phonetics and phonology, their knowledge continues to develop throughout childhood (e.g., CITE). One common task in this line of research is nonword repetition (NWR). In NWR studies, participants are presented auditorily with an item that is phonologically legal but lexically meaningless in the language children are learning. The child should immediately try to say it back without changing anything. Accuracy is thought to reflect long-term phonological knowledge (which allows the child to perceive the item accurately even though it is not a real word they have encountered before) as well as online working memory (to encode the item in the interval between hearing it and saying it back).

NWR has been used to seek answers to theoretical questions, including what are the links between phonology, working memory, and the lexicon {[}CITE{]}, as well as for applications, notably using nwr as a diagnostic for language delays and disorders {[}CITE{]}. Since non-words can be generated in many languages, it has been used acros a wide array of languages, particularly in Europe {[}CITE{]}.

@@

Although the procedure tends to be quite stable, there is some variation in the presentation and structure of items found in previous NWR work. For example, items are often presented orally by the experimenter {[}CITE{]}, although an increasing number of studies have turned to playing back the stimuli in order to have greater control of the stability of the presentation {[}CITE{]}. Additionally, while some studies have used 10-15 non-words, others have employed up to 30 unique items {[}CITE{]}. Often, authors modulate structural complexity, typically measured in terms of item length (measured in number of syllables) as well as syllable structure {[}open as opposed to closed syllables, CITE{]}.

@@

Naturally, designing the task in this way may render the study of individual variaiton within the population more difficult because different children are exposed to different items. However, a review of previous work on individual variation suggested to us that many individual differences effects are relatively small, and would not be detectable with the sample size that we could collect in a given visit. That said, we contribute to the literature by also reporting descriptive analyses that could potentially be integrated in meta-analytic efforts. Based on previous work, we looked at potential improvements with age {[}CITE{]}, and potential negative effects of bilingual exposure {[}CITE{]}. Previous work typically finds no significant differences as a function of maternal education {[}see CITE but also CITE{]} or child gender {[}CITE{]}. Although previous research has not often investigated potential effects of birth order on NWR, there is a sizable literature on these effects in other language tasks {[}CITE{]}, and therefore we report on those too.

@@

\begin{itemize}
\tightlist
\item
  What is the overall repetition accuracy (whole word, phoneme based, distance)?
\item
  How does this change as a function of item complexity (number of syllablex, sound complexity)?
\item
  How frequent are errors that result in real words? Is that a function of item complexity? What are some common patterns of phonological errors?
\item
  Is individual variation explainable by child age, sex, birth order, monolingual status, and/or parental education?
\end{itemize}

@@

({\textbf{???}}) ADD TABLE SUMMARIZING PREVIOUS WORK

\begin{itemize}
\tightlist
\item
  age differences
\item
  length effects
\end{itemize}

@@

A second list of candidate items was generated in 2019 by selecting complex consonants and systematically crossing them with all the vowels in the Yelî inventory to produce consonant-vowel monosyllables. As before, items were automatically excluded if they appeared in the dictionary. Additionally, since hearing vowel length in monosyllables in isolation is challenging, any item that had a short/long real word neighbor was filtered out. Since the phonology and phonetics of Yélî is still in the process of being described {[}CITE{]}, there could have been undocumented constraints that rendered items illegal. Therefore, we made sure that the precise consonant-vowel sequence occurred in some real word in the dictionary (i.e., that there was a longer word included the monosyllable as a subsequence). These candidates were presented to one informant, for a final check that they did not mean anything. Together with the 2018 selection, they were recorded using a headset XXX and an olympus XXX from the written form presented together with the same item orally. The complete recorded list was finally presented to two more informants, who could repeat all the items and who confirmed there were no real words. Even so, there was one monosyllable that was often identified as a real word (intended \enquote{yî} /yXX/; identified as \enquote{yi} /yi/, \emph{tree}). This item is removed from analyses.

@@

The child was donned a headset (xx ({\textbf{???}}) please fill in for most of the children, xx ({\textbf{???}}) fill in for the rest). For most children, the headset could not stay comfortably on the child's head, and thus it was placed on the child's shoulders, with the microphone carefully placed close to the child's mouth. A local informant sat next to the child, to would provide the instructions and, if needed, coach the child to make sure, using the three practice items as well as real words, that they understood that the task was to repeat the items precisely without changing anything. An experimenter delivered the elicitation stimuli to the local informant and the child over headhphones.

@@

\begin{itemize}
\item
  What is the overall repetition accuracy (whole word, phoneme based, distance)?
\item
  How does this change as a function of item complexity (number of syllablex, sound complexity)?
\item
  How frequent are errors that result in real words? Is that a function of item complexity? What are some common patterns of phonological errors?
\item
  Is individual variation explainable by child age, sex, birth order, monolingual status, and/or parental education?
\end{itemize}

\begin{verbatim}
##     ID score.online
## 1  c70    0.7307692
## 2  c71    0.5384615
## 3  c72    0.6086957
## 4  c73    0.8461538
## 5  c74    0.9090909
## 6  c76    0.7307692
## 7  c77    0.9565217
## 8  c78    0.8461538
## 9  c79    0.7692308
## 10 c80    0.8260870
## 11 c81    0.5652174
## 12 c82    0.5000000
\end{verbatim}

\includegraphics{manuscript_files/figure-latex/unnamed-chunk-1-1.pdf} \includegraphics{manuscript_files/figure-latex/unnamed-chunk-1-2.pdf} \includegraphics{manuscript_files/figure-latex/unnamed-chunk-1-3.pdf}

@@

\begin{itemize}
\tightlist
\item
  What is the overall repetition accuracy (whole word, phoneme based, distance)?
\item
  How does this change as a function of item complexity (number of syllablex, sound complexity)?
\item
  How frequent are errors that result in real words? Is that a function of item complexity?
\item
  Is individual variation explainable by child age, sex, birth order, monolingual status, and/or parental education?
\end{itemize}

@@

>>>>

\hypertarget{todo}{%
\subsection{TODO}\label{todo}}

\begin{itemize}
\tightlist
\item
  create stimuli table that has complexity \& length
\item
  integrate demo data \& stim table in final data
\item
  add analysis on the basis of complexity \& length
\item
  add analysis on the basis of demo chars
\item
  fix the back-transformation
\item
  add summary of types of error: \%semantic, \%deletion (which sounds?) \%subst (which are the most common sounds)
\item
  maybe add to the last ana the prevalence (ie out of all the gh which proportion get transformed)
\item
  discussion
\end{itemize}

@@

Although infants begin to learn about their native language's phonology within the first year, many studies suggest that in perception and production, in phonetics and phonology, their knowledge continues to develop throughout childhood (e.g., Hazan \& Barrett, 2000). One common task in this line of research is nonword repetition (NWR). In NWR studies, participants are presented auditorily with an item that is phonologically legal but lexically meaningless in the language children are learning. The child should immediately try to say it back without changing anything. Accuracy is thought to reflect long-term phonological knowledge (which allows the child to perceive the item accurately even though it is not a real word they have encountered before) as well as online working memory (to encode the item in the interval between hearing it and saying it back).

NWR has been used to seek answers to theoretical questions, including what are the links between phonology, working memory, and the lexicon (Bowey, 2001), as well as for applications, notably using nwr as a diagnostic for language delays and disorders (Estes, Evans, \& Else-Quest, 2007). Since non-words can be generated in many languages, it has been used acros a wide array of languages, particularly in Europe (Meir, Walters, \& Armon-Lotem, 2016,@is08042009language).

@@

Although the procedure tends to be quite stable, there is some variation in the presentation and structure of items found in previous NWR work. For example, items are often presented orally by the experimenter (Torrington Eaton, Newman, Ratner, \& Rowe, 2015), although an increasing number of studies have turned to playing back the stimuli in order to have greater control of the stability of the presentation (Brandeker \& Thordardottir, 2015). Additionally, while some studies have used 10-15 non-words, others have employed up to 46 unique items (Piazzalunga, Previtali, Pozzoli, Scarponi, \& Schindler, 2019). Often, authors modulate structural complexity, typically measured in terms of item length (measured in number of syllables) and/or syllable structure (open as opposed to closed syllables, Gallon, Harris, \& Van der Lely, 2007).

@@

Naturally, designing the task in this way may render the study of individual variaiton within the population more difficult because different children are exposed to different items. However, a review of previous work on individual variation suggested to us that many individual differences effects are relatively small, and would not be detectable with the sample size that we could collect in a given visit. That said, we contribute to the literature by also reporting descriptive analyses that could potentially be integrated in meta-analytic efforts. Based on previous work, we looked at potential improvements with age (Farmani et al., 2018; Kalnak, Peyrard-Janvid, Forssberg, \& Sahlén, 2014; Vance, Stackhouse, \& Wells, 2005), and potential negative effects of bilingual exposure (Brandeker \& Thordardottir, 2015; Meir \& Armon-Lotem, 2017; Meir et al., 2016). Previous work typically finds no significant differences as a function of maternal education (e.g., Farmani et al., 2018; Kalnak et al., 2014; Meir \& Armon-Lotem, 2017) or child gender (Chiat \& Roy, 2007). Although previous research has not often investigated potential effects of birth order on NWR, there is a sizable literature on these effects in other language tasks (Havron et al., 2019), and therefore we report on those too.

@@

\begin{itemize}
\tightlist
\item
  What is the overall repetition accuracy (whole word, phoneme based, distance)?
\item
  How does this change as a function of item complexity (number of syllables, sound complexity)?
\item
  How frequent are errors that result in real words? Is that a function of item complexity? What are some common patterns of phonological errors?
\item
  Is individual variation explainable by child age, sex, birth order, monolingual status, and/or parental education?
\end{itemize}

@@

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/prevlit-fig-1.pdf}
\caption{\label{fig:prevlit-fig}Proportion of non-words correctly repeated as a function of age (in years), study (first author and year regardless of the number of authors), and, when available, shorter versus longer non-words.}
\end{figure}

@@

A second list of candidate items was generated in 2019 by selecting complex consonants and systematically crossing them with all the vowels in the Yelî inventory to produce consonant-vowel monosyllables. As before, items were automatically excluded if they appeared in the dictionary. Additionally, since hearing vowel length in monosyllables in isolation is challenging, any item that had a short/long real word neighbor was filtered out. Since the phonology and phonetics of Yélî is still in the process of being described (CITE {\textbf{???}} please fill in), there could have been undocumented constraints that rendered items illegal. Therefore, we made sure that the precise consonant-vowel sequence occurred in some real word in the dictionary (i.e., that there was a longer word included the monosyllable as a subsequence). These candidates were presented to one informant, for a final check that they did not mean anything. Together with the 2018 selection, they were recorded using a headset XXX and an olympus XXX from the written form presented together with the same item orally. The complete recorded list was finally presented to two more informants, who could repeat all the items and who confirmed there were no real words. Even so, there was one monosyllable that was often identified as a real word (intended \enquote{yî} /yXX/; identified as \enquote{yi} /yi/, \emph{tree}). This item is removed from analyses.

@@

The child was donned a headset (xx ({\textbf{???}}) please fill in for most of the children, SHURE WH20 XLR headset with a dynamic microphone for the rest), recorded into the left channel into a Tascam DR40x digital recorder. For most children, the headset could not stay comfortably on the child's head, and thus it was placed on the child's shoulders, with the microphone carefully placed close to the child's mouth. A local informant sat next to the child, to would provide the instructions and, if needed, coach the child to make sure, using the three practice items as well as real words, that they understood that the task was to repeat the items precisely without changing anything. An experimenter delivered the elicitation stimuli to the local informant and the child over headphones.

@@

Taking into account only the first attempts, the overall repetition accuracy measured in whole words is 73\% based on the online coding (i.e., whether the native experimenter asked for the target to be re-played), and 65\% when the offline coding is considered. On average, the phoneme-based normalized Levenshtein distance was 20\%, meaning that about 20\% of phonemes were substituted, inserted, or deleted.

We then inspected whether accuracy varied as a function of word length.

\begin{verbatim}
##     Group.1 Group.2 x
## 1         2    c105 0
## 2         3    c105 0
## 3         4    c105 0
## 4         6    c105 0
## 5         8    c105 0
## 6         3    c106 0
## 7         4    c106 0
## 8         6    c106 0
## 9         2    c107 0
## 10        3    c107 0
## 11        4    c107 0
## 12        6    c107 0
## 13        3    c108 0
## 14        4    c108 0
## 15        6    c108 0
## 16        8    c108 0
## 17        2    c109 0
## 18        3    c109 0
## 19        4    c109 0
## 20        6    c109 0
## 21        3    c110 0
## 22        4    c110 0
## 23        6    c110 0
## 24        8    c110 0
## 25        3    c111 0
## 26        4    c111 0
## 27        8    c111 0
## 28        2    c112 0
## 29        3    c112 0
## 30        4    c112 0
## 31        3    c114 0
## 32        4    c114 0
## 33        6    c114 0
## 34        8    c114 0
## 35        3    c115 0
## 36        4    c115 0
## 37        6    c115 0
## 38        8    c115 0
## 39        3    c116 0
## 40        6    c116 0
## 41        8    c116 0
## 42        3    c117 0
## 43        4    c117 0
## 44        3    c118 0
## 45        4    c118 0
## 46        6    c118 0
## 47        8    c118 0
## 48        3    c119 0
## 49        6    c119 0
## 50        8    c119 0
## 51        3    c120 0
## 52        4    c120 0
## 53        6    c120 0
## 54        8    c120 0
## 55        3    c121 0
## 56        4    c121 0
## 57        3    c122 0
## 58        4    c122 0
## 59        6    c122 0
## 60        8    c122 0
## 61        3    c123 0
## 62        4    c123 0
## 63        2     c70 0
## 64        3     c70 0
## 65        4     c70 0
## 66        6     c70 0
## 67        8     c70 0
## 68        2     c71 0
## 69        3     c71 0
## 70        4     c71 0
## 71        6     c71 0
## 72        8     c71 0
## 73        3     c72 0
## 74        4     c72 0
## 75        6     c72 0
## 76        8     c72 0
## 77        3     c73 0
## 78        4     c73 0
## 79        6     c73 0
## 80        8     c73 0
## 81        2     c74 0
## 82        3     c74 0
## 83        4     c74 0
## 84        6     c74 0
## 85        8     c74 0
## 86        3     c76 0
## 87        4     c76 0
## 88        8     c76 0
## 89        3     c77 0
## 90        4     c77 0
## 91        6     c77 0
## 92        8     c77 0
## 93        2     c78 0
## 94        3     c78 0
## 95        8     c78 0
## 96        2     c79 0
## 97        3     c79 0
## 98        2     c80 0
## 99        3     c80 0
## 100       4     c80 0
## 101       6     c80 0
## 102       8     c80 0
## 103       2     c81 0
## 104       3     c81 0
## 105       4     c81 0
## 106       8     c81 0
## 107       3     c82 0
## 108       4     c82 0
## 109       6     c82 0
## 110       8     c82 0
## 111       3     c83 0
## 112       4     c83 0
## 113       6     c83 0
## 114       8     c83 0
## 115       3     c84 0
## 116       4     c84 0
## 117       6     c84 0
## 118       8     c84 0
## 119       3     c85 0
## 120       4     c85 0
## 121       6     c85 0
## 122       8     c85 0
## 123       3     c87 0
## 124       4     c87 0
## 125       6     c87 0
## 126       8     c87 0
## 127       3     c88 0
## 128       4     c88 0
## 129       3     c89 0
## 130       4     c89 0
## 131       6     c89 0
## 132       3     c90 0
## 133       4     c90 0
## 134       6     c90 0
## 135       8     c90 0
## 136       2     c91 0
## 137       3     c91 0
## 138       4     c91 0
## 139       2     c93 0
## 140       3     c93 0
## 141       4     c93 0
## 142       6     c93 0
## 143       3     c94 0
## 144       4     c94 0
## 145       8     c94 0
## 146       3     c95 0
## 147       4     c95 0
## 148       3     c99 0
## 149       4     c99 0
## 150       6     c99 0
\end{verbatim}

\includegraphics{manuscript_files/figure-latex/figlength-1.pdf}
- How frequent are errors that result in real words? Is that a function of item complexity? What are some common patterns of phonological errors?

\begin{itemize}
\tightlist
\item
  Is individual variation explainable by child age, sex, birth order, monolingual status, and/or parental education?
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  What is the overall repetition accuracy (whole word, phoneme based, distance)?
\item
  How does this change as a function of item complexity (number of syllables, sound complexity)?
\item
  How frequent are errors that result in real words? Is that a function of item complexity?
\item
  Is individual variation explainable by child age, sex, birth order, monolingual status, and/or parental education?
\end{itemize}

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-bowey2001nonword}{}%
Bowey, J. A. (2001). Nonword repetition and young children's receptive vocabulary: A longitudinal study. \emph{Applied Psycholinguistics}, \emph{22}(3), 441--469.

\leavevmode\hypertarget{ref-brandeker2015language}{}%
Brandeker, M., \& Thordardottir, E. (2015). Language exposure in bilingual toddlers: Performance on nonword repetition and lexical tasks. \emph{American Journal of Speech-Language Pathology}, \emph{24}(2), 126--138.

\leavevmode\hypertarget{ref-chiat2007preschool}{}%
Chiat, S., \& Roy, P. (2007). The preschool repetition test: An evaluation of performance in typically developing and clinically referred children. \emph{Journal of Speech, Language, and Hearing Research}, \emph{50}(2), 429--443.

\leavevmode\hypertarget{ref-estes2007differences}{}%
Estes, K. G., Evans, J. L., \& Else-Quest, N. M. (2007). Differences in the nonword repetition performance of children with and without specific language impairment: A meta-analysis. \emph{Journal of Speech, Language, and Hearing Research}, \emph{50}(1), 177--195.

\leavevmode\hypertarget{ref-farmani2018normalization}{}%
Farmani, H., Sayyahi, F., Soleymani, Z., Labbaf, F. Z., Talebi, E., \& Shourvazi, Z. (2018). Normalization of the non-word repetition test in farsi-speaking children. \emph{Journal of Modern Rehabilitation}, \emph{12}(4), 217--224.

\leavevmode\hypertarget{ref-gallon2007non}{}%
Gallon, N., Harris, J., \& Van der Lely, H. (2007). Non-word repetition: An investigation of phonological complexity in children with grammatical sli. \emph{Clinical Linguistics \& Phonetics}, \emph{21}(6), 435--455.

\leavevmode\hypertarget{ref-havron2019effect}{}%
Havron, N., Ramus, F., Heude, B., Forhan, A., Cristia, A., Peyre, H., \& Group, E. M.-C. C. S. (2019). The effect of older siblings on language development as a function of age difference and sex. \emph{Psychological Science}, \emph{30}(9), 1333--1343.

\leavevmode\hypertarget{ref-hazan2000development}{}%
Hazan, V., \& Barrett, S. (2000). The development of phonemic categorization in children aged 6--12. \emph{Journal of Phonetics}, \emph{28}(4), 377--396.

\leavevmode\hypertarget{ref-is08042009language}{}%
IS0804, C. A. (2009). Language impairment in a multilingual society: Linguistic patterns and the road to assessment. \emph{Brussels: COST Office. Available Online at: Http://Www. Bi-Sli. Org}.

\leavevmode\hypertarget{ref-kalnak2014nonword}{}%
Kalnak, N., Peyrard-Janvid, M., Forssberg, H., \& Sahlén, B. (2014). Nonword repetition--a clinical marker for specific language impairment in swedish associated with parents' language-related problems. \emph{PloS One}, \emph{9}(2), e89544.

\leavevmode\hypertarget{ref-meir2017independent}{}%
Meir, N., \& Armon-Lotem, S. (2017). Independent and combined effects of socioeconomic status (ses) and bilingualism on children's vocabulary and verbal short-term memory. \emph{Frontiers in Psychology}, \emph{8}, 1442.

\leavevmode\hypertarget{ref-meir2016disentangling}{}%
Meir, N., Walters, J., \& Armon-Lotem, S. (2016). Disentangling sli and bilingualism using sentence repetition tasks: The impact of l1 and l2 properties. \emph{International Journal of Bilingualism}, \emph{20}(4), 421--452.

\leavevmode\hypertarget{ref-piazzalunga2019articulatory}{}%
Piazzalunga, S., Previtali, L., Pozzoli, R., Scarponi, L., \& Schindler, A. (2019). An articulatory-based disyllabic and trisyllabic non-word repetition test: Reliability and validity in italian 3-to 7-year-old children. \emph{Clinical Linguistics \& Phonetics}, \emph{33}(5), 437--456.

\leavevmode\hypertarget{ref-torrington2015non}{}%
Torrington Eaton, C., Newman, R. S., Ratner, N. B., \& Rowe, M. L. (2015). Non-word repetition in 2-year-olds: Replication of an adapted paradigm and a useful methodological extension. \emph{Clinical Linguistics \& Phonetics}, \emph{29}(7), 523--535.

\leavevmode\hypertarget{ref-vance2005speech}{}%
Vance, M., Stackhouse, J., \& Wells, B. (2005). Speech-production skills in children aged 3--7 years. \emph{International Journal of Language \& Communication Disorders}, \emph{40}(1), 29--48.
